---
title: "Reproducible Research Assignment 1"
author: "MGreels"
date: "8/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Reproducible Research - Programming Assignment #1

## Loading and Preprocessing the Data into R

We start by downloading the dataset and saving to an R dataframe.  

The raw dataframe is also reconfigured into two aggregate datasets, one taking a daily total step count sum for all the days in the study and the second computing an average for each 5 minute interval across the study set.

```{r load, echo=TRUE}
RawDat <- read.csv("activity/activity.csv")


DayAves <- aggregate(RawDat$steps, 
                     list(Day = RawDat$date), 
                     sum,
                     na.rm = TRUE)

IntAves <- aggregate(RawDat$steps, 
                     list(Interval = RawDat$interval), 
                     mean, 
                     na.rm = TRUE)
```

## What is the Mean total Number of Steps taken per Day?

```{r Mean total Steps}
dailymean <- mean(DayAves$x)
dailymed <- median(DayAves$x)

library(ggplot2)

histo <- ggplot(data=subset(DayAves, !is.na(x)), aes(x)) + 
        geom_histogram(bins=10)

histo

```

The mean total daily steps recorded was `r dailymean` and teh median total daily steps recorded was `r dailymed`

## What is the Average Daily Activity Pattern?

```{r Daily Timestamp Averages}
library(ggplot2)

maxint  <- IntAves$Interval[which.max(IntAves$x)]

barplot <- ggplot(data=IntAves, aes(x = Interval, y = x)) +
                geom_bar(stat = "identity")

barplot
```

The maximum steps occur daily during interval number `r maxint`.

## Impute Missing Values

Code Below will create a new dataset "ImpDat" with imputed data to replace NAs.  In the case of an NA, the code inserts the associated 5 minute interval average.

Code then reruns Daily and per interval averages for the two dataframes created above.

```{r Impute Missing Values}
ImpDat<- RawDat
for (i in 1:length(RawDat$steps)){
        if (is.na(ImpDat[i,1])){
                ImpDat[i,1] <- IntAves$x[
                        which(IntAves$Interval == round(ImpDat[i,3]))] 
        }
}

ImpDA <- aggregate(ImpDat$steps, 
                     list(Day = ImpDat$date), 
                     sum)

ImpIA <- aggregate(ImpDat$steps, 
                     list(Interval = ImpDat$interval), 
                     mean)

impmean <- mean(ImpDA$x)
impmed <- median(ImpDA$x)

imphist <- ggplot(data=subset(ImpDA, !is.na(x)), aes(x)) + 
        geom_histogram(bins=10)

imphist


```

WIth missing data imputed the new daily mean step count is `r impmean` and the new daily median step count is `r impmed`


## Are there differences in activity patterns between weekdays and weekends?

```{r Weekend v. Weekday Analysis}
library(chron)

ImpIAWE <- aggregate(ImpDat$steps, 
                     list(Interval = ImpDat$interval, 
                          weekend = is.weekend(as.Date(ImpDat$date))), 
                     mean)

ImpIAWE$WE <- factor(ImpIAWE$weekend, labels = c("Weekday", "Weekend"))

WEplot <- ggplot(data=ImpIAWE, aes(x = Interval, y = x)) +
        geom_bar(stat = "identity") +
        facet_grid(WE~.,)

WEplot
```